Wikipedia Text Mining and Entropy Analysis

### 2019-01-22 cogsci/position-info.Rmd
updated CogSci intro, working on clustering comparison

---
title: "A Language's Unigram Entropy Distribution Predicts Self-Paced Reading Times [Rewrite title]"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

How do people convey information in speech and text? Is there a regular distribution to how information is conveyed across speech and text? Does speech, with its phonological properties, different from writing, with its semiotic properties, in how information is conveyed? [Are we pitting speaking against writing?]

A major line of research in the current century has asked how speakers and writers distribute information in the linguistic material they produce. This begins with Claude Shannon in the years after World War II. Shannon (1948) defined "information" as a "reduction in uncertainty". Shannon followed by quantifying uncertainty in his measure of *entropy*: the amount of uncertainty on the outcome of a random variable. Shannon proposed that the most efficient method of sending information through a noisy channel is at a constant rate. Genzel & Charniak (2002) apply this principle to human communication: if people communicate information through speech and text optimally, then we should transmit information to one another at a constant rate. Genzel & Charniak created a distribution of estimates for sentence-level entropy, each sentence taken without context, for newspaper articles from the Penn Treebank corpus. Obtaining a roughly monotonically increasing linear distribution, Genzel & Charniak proposed the *Constant Entropy Rate* (CER) principle: the entropy rate in speech and text with context should be constant as sentence number increases. Aylett & Turk (2004) proposed a similar principle for the phonological level.

Jaeger (2010) extends the CER principle to all levels of human communication: the *Uniform Information Density* (UID) hypothesis. Jaeger argues that speakers will try to distribute information more evenly over the course of an utterance, to be as close to channel capacity (in a Shannon sense) as possible. In particular, Jaeger singles out the production of "that" at the beginning of relative clauses in English and the use of contractions such as "he's" and "you're" versus "he is" and "you are" as instances where higher information density in the clause elicits the production of more material by speakers [Explain the Jaeger data more clearly and succintly, connect with other examples in the literature more broadly].

More recent work has stood out in contrast to the traditional UID perspective. Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifiers, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by UID. The UID perspective is challenged in Yu et al. (2016), by performing an analysis of entropy by position within sentence in the text portion of the British National Corpus. This work serves as the basis for ours.

@@

To explore information transmission rate in speech, we used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations [Why does it matter to examine child vs adult data]. We first used the Brown and Providence English corpora from CHILDES. The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided each corpus by speaker into child and non-child categories. We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position. We follow Yu et al. (2016) in using the following formula for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e. the number of sentences of length $k$. [Give the equation its own paragraph]

The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances in the English CHILDES corpora. The entropy curves capture individual variation across positions in utterances of the same length. This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence. For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora. We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances. We are applying the same approach as in Genzel and Charniak (2003), but within sentences instead of across sentences.

@@

# Results & Analysis
The Unigram Information Density account would predict that children have a lower channel capacity for communication and would instead communicate with a less unigram entropy curve. The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce . This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

>>>>

---
title: "Rate of Information "
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

When people communicate in speaking or writing, they can have a variety of different goals. Perhaps the most important of these goals is to communicate information. If we take this as the primary goal of communication, then how information is transmitted should have regular properties. There is a rich literature on information, dating back to Claude Shannon’s theory of information (1948). We know from Shannon’s work that rate of information transmission should be constant and as close as possible to the capacity of the communication channel. Levy & Jaeger (2007) introduce the *uniform information density* hypothesis, arguing that speakers structure utterances so as to optimize information density. Viewing human communication as a noisy channel, the authors argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible. Their evidence comes in the form of the use or lack thereof of “that”-complementizers in introducing English relative clauses. When the word preceding the relative clause or the initial word of the relative clause has high information content, then the authors argue that speakers are more likely to use the complementizer. The authors use adult-adult conversations from the Penn Treebank corpus (CITATION).

Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifiers, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by the UID perspective.

Let us assume that the UID perspective is correct. Then entropy should monotonically and uniformly increase at all word positions across utterances on average. We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.

In Yu et al. (2016), the authors challenge the UID perspective through computing the unigram positional entropy within the text portion of the British National Corpus (BNC). We follow Yu et al. (2016) in using the following formula for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e. the number of sentences of length $k$. [Give the equation its own paragraph]

The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances

@@

For spoken data, we choose to examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages. We expect based on the UID account that the entropy curves of the children will be significantly lower than those of the adults, due to the children's comparatively lower channel capacity in information transmission [more support]. We first used the Brown and Providence English corpora from CHILDES. The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided each corpus by speaker into child and non-child categories. We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position.

In the English CHILDES corpora. The entropy curves capture individual variation across positions in utterances of the same length. This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence. For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora. We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances. We are applying the same approach as in Genzel and Charniak (2003), but within sentences instead of across sentences.

@@

# Results & Analysis
The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce . This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

### 2019-01-22 wiki_entropy/wiki_clustering.Rmd
updated CogSci intro, working on clustering comparison

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lingtypology)
library(knitr)
library(klaR)
library(missMDA)
library(clusteval)
library(factoextra)
library(cluster)
library(widyr)

@@

```{r Gelman}
## from http://www.stat.columbia.edu/~gelman/bda/regression.R

sigma.hat <- function (object){
  object.class <- class(object)[[1]]
  if (object.class=="lm"){
    sigma <- summary(object)$sigma
    return (sigma)
  }
  else if (object.class=="glm"){
    sigma <- summary(object, correlation=TRUE)$sigma
    return (sigma)
  }
  else if (object.class=="lmer"){
    object <- summary (object)
    fcoef <- .Call("mer_fixef", object, PACKAGE = "lme4")
    useScale <- object@devComp[8]
    ngrps <- lapply(object@flist, function(x) length(levels(x)))
    n.groupings <- length (ngrps)
    varc <- VarCorr (object, useScale=useScale)
    sc <- attr(varc, "sc")
    recorr <- lapply(varc, function(el) el@factors$correlation)
    reStdDev <- c(lapply(recorr, slot, "sd"), list(Residual = sc))
    sigmas <- as.list (rep (NA, n.groupings+1))
    sigmas[1] <- ifelse (useScale, sc, NA)
    cors <- as.list (rep (NA, n.groupings+1))
    names (sigmas) <- names (cors) <- c ("data", names (varc))
    for (k in 1:n.groupings){
      sigmas[[k+1]] <- reStdDev[[k]]
      cors[[k+1]] <- as.matrix (recorr[[k]])
      if (length (cors[[k+1]]) == 1) cors[[k+1]] <- NA
    }
    return (list (sigma=sigmas, cors=cors))
  }
}
```

```{r WALS_imputation_functions}
#taken from https://www.tutorialspoint.com/r/r_mean_median_mode.htm
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

## imputation based on Gelman & Hill Chapter 25
random.imp <- function(a) {
  missed <- is.na(a)
  num_missed <- sum(missed)
  a.obs <- a[!missed]
  imputed <- a
  imputed[missed] <- base::sample(a.obs, num_missed, replace=TRUE)
  return(imputed)
}

impute <- function(a, a.impute) {
  ifelse(is.na(a), a.impute, a)
}

get_factor <- function(cname) {
  a.imp <- wf.sub[c(cname)] %>%
    random.imp() %>%
    unlist(use.names = FALSE)
}
```

```{r WALS_imputation_code}
FEATURES <- c(1:144)
for (i in c(1:144)) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

NUM_CLUSTERS <- 4

wf <- wals.feature(FEATURES)
wf.sub <- wf[c("1A", "2A", "3A", "4A", "language")]
wf.sub <- wf.sub %>% 
  filter(language %in% LANGUAGES) %>%
  unique()

a1.imp <- get_factor("1A")
a2.imp <- get_factor("2A")
a3.imp <- get_factor("3A")
a4.imp <- get_factor("4A")

a1.mode <- getmode(wf.sub$`1A`[!is.na(wf.sub$`1A`)]) #gets the most common non-na value of column 1A

n.sims <- 10
for (s in 1:n.sims) {
  a1 <- wf.sub$`1A`
  a1[is.na(a1)] <- getmode(a1[!is.na(a1)])
  lm.1 <- glm(a1 ~ a2.imp + a3.imp + a4.imp, family = "binomial")
  pred.1 <- rnorm(n, predict(lm.1), sigma.hat(lm.1)) ####
  a1.imp <- impute(wf.sub[c("1A")], pred.1)
}

```{r WALS_clustering}
FEATURES <- c(1:144)
for (i in c(1:144)) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

wf <- wals.feature(FEATURES)
wf %>% View()
wf <- wf[, -(1:3)]
wf <- wf[, -(ncol(wf) - 1)]
wf <- wf %>% 
  filter(language %in% LANGUAGES)
wf <- wf[, complete.cases(wf)] #this significantly trims down the dataset but is necessary for k-modes
rv <- kmodes(wf[, -(ncol(wf))], NUM_CLUSTERS, iter.max=10, weighted = FALSE)

@@

wf_sub <- wf_df %>%
  filter(feature %in% wf_many_langs$feature,
         language %in% wf_sub_langs$language) %>%
  select(language, feature, value) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value))) %>%
  group_by(feature) %>% # for each pair of language, does this feature assin them the same value
  nest() %>%
  mutate(cors = map(data, my_pairwise_cor))
  select(-data) %>%
  unnest()

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lingtypology)
library(knitr)
library(missMDA)
library(cluster)
library(widyr)

@@

@@

wf_sub <- wf_df %>%
  filter(feature %in% wf_many_langs$feature,
         language %in% wf_sub_langs$language) %>%
  dplyr::select(language, feature, value) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value))) %>%
  group_by(feature) %>% # for each pair of language, does this feature assin them the same value
  nest() %>% #maps features to a table mapping languages to numerical factor feature values
  mutate(cors = map(data, my_pairwise_cor))
  select(-data) %>%
  unnest()

wf_comp <- wf_sub %$%
  data %>%
  do.call(cbind, args=.) 
wf_comp <- wf_comp[, -seq(3, length(wf_comp), 2)] #get values and one column of language

langs_to_compare <- wf_comp$language
wkpd_comp <- wkpd %>%
    filter(Language %in% langs_to_compare)
